---
title: "R Tools For Course Management - Development Documentation"
author: "Peter von Rohr"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{R Tools For Course Management}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r Prereq, echo=FALSE, results="hide"}
rm(list = ls())
if (!require(devtools)) install.packages("devtools")
if (!require(stringi)) install.packages("stringi")
if (!require(knitr)) install.packages("knitr")
if (!require(slidify)) devtools::install_git(c("ramnathv/slidify", "ramnathv/slidifyLibraries"))
#sCourseToolsPackDir <- ifelse(.Platform$OS.type == "windows", 
#                              "c:/Daten/pvr/Google Drive/Projects/RStudioProjects/RCourseTools", 
#                             "/Users/peter/Google Drive/Projects/RStudioProjects/RCourseTools")
devtools::load_all()
```

## Introduction
Course management involves a standard set of tasks. Once a course is planned, outlined and scheduled, the following tasks must to be executed.

* Deciding on course topics
* Distributing topics across lecture dates
* Writing course syllabus
* Collecting material for the course
* Writing course notes
* Putting together course slides
* Preparing lectures
* Assigning homework to students
* Collecting, grading and returning students homework

Support for course management can be provided by a set of tools which fascilitate the above listed tasks.

### Available Course Tools
[SageMathCloud](http://cloud.sagemath.org) (SMC) is an open source platform which provides extensive functionality and already solves many of the above tasks. The only thing that is missing so far is convenient offline support. That means as soon as one is disconnected from the internet, it is impossible to use the tools provided by SMC. There is always the possibility to download the files that one is working on to the local machine before going offline and uploading them again when getting re-connected, but that is somewhat tedious and most users would probably wish for a more seamless workflow.  

Secondly, when working with the course tools on SMC, all participants of the course such as co-teachers, TAs and students are forced to have an account on SMC. While this might be fine for most course participants, there might be some persons who refuse to create an account on SMC. Furthermore, once course participants have an account, they are required to remember their account credentials which is also not a problem for most participants, but it might be painful for the two or three students who always forget their password and have to get it reset all the time. These problems can be addressed by distributing the course material via a git repository such as Github. Combining the distribution of course material via Github together with R packages such as `devtools`, it is very convenient to clone Github repositories. Furthermore R provides a convenient packaging infrastructure which can be used to organize the complete course material.  


## A Course Model
When trying to come up with a model for a course, one has to distinguish between at least two views. 

1. There is the content or topic oriented view. Such types of models have been developed in Michaela Pedronis [PhD dissertation](http://e-collection.library.ethz.ch/eserv/eth:1295/eth-1295-02.pdf). The main result of that dissertation is that courses can be modelled using so called `testable and reusable components` (TRUC).
2. The second view is more focused on the representation of the different components of a course. Here the components are course notes, slides, course syllabi, course websites, etc. Those components can all easily be modelled by very generic objects such as URIs, documents or strings. 

The main focus of this package is to automate and to fascilitate as much as possible the operations with the components of the representation view of a course.


## The Plan 
The goal is to have the same functionality as other open source online course tools such as SMC and to have a more seamless workflow and better offline support provided by the set of tools. 

This goal is achieved by combining the R packaging mechanism together with the Github support provided by certain R-packages such as `devtools` and by R-Studio. More concretely, the complete course material is contained in an R package which can easily be downloaded and installed using the R package `devtools`. The main advantage of this approach of putting all course material into an R package is that apart from R, RStudio and `devtools` which are required anyway for the course, no additional software needs to be installed or no account must be created by the students.

### Proposed workflow
When starting from scratch, the first thing is to come up with an initial skeleton for a course. This can be done by a call to the function `create_course()`. The only mandatory argument that has to be specified when calling `create_course()` is the name of the course that we want to create. The course skeleton that is created by `create_course()` consists of a number of subdirectories and files all grouped in a directory that has the same name as the course. When not specifying any directory argument to `create_course` then the current working directory will be used as parent directory for the course. 

### A first attempt
The following code section shows a first version of a function that creates a course skeleton using `devtools::create()` to come up with an initial course package

```{r CreateCourseFirstVersion}
#' Create a course with a given name under a given working directory
#'
#' \code{create_course} uses devtools::create to create a package skeleton
#'
#' @param psCourseName      name of the course
#' @param psCourseWorkDir   course working directory
#' @example create_course("MyCourse")
#' @export
create_course <- function(psCourseName, psCourseWorkDir = getwd()) {
  ##
  ##   create_course(psCourseName, psCourseWorkdir): creates the 
  ##      skeleton of a course in the form of an R package
  ## ########################################################### ##
  ## # check that at least course name argument is specified
  stopifnot(!is.null(psCourseName))
  ## # if the parent directory of the course directory does not exist, create it
  if (! dir.exists(psCourseWorkDir)) dir.create(psCourseWorkDir)
  ## # create an R package using devtools::create
  sCoursePath <- file.path(psCourseWorkDir,psCourseName)
  ## # in case old versions exist, remove them
  if (dir.exists(sCoursePath)) unlink(sCoursePath, recursive = TRUE, force = TRUE)
  ## # use devtools to create the package
  devtools::create(path = sCoursePath)
  cat("\n * Created course: ", psCourseName, " in directory: ", psCourseWorkDir, "\n")
  invisible()
}
```

Now that we have created a first version of a function that creates a course skeleton, we are ready to test it. Before the call to `create_course`, we are assigning the course name, the course working directory and the course description to variables. 


```{r}
sCourseName <- "rintrocourse"
sCourseWorkDir <- ifelse(.Platform$OS.type == "windows", 
                         "c:/Daten/pvr/Projects/RStudioProjects/Courses", 
                         "/Users/peter/Data/Projects/RStudioProjects/Courses")
sCoursePath <- file.path(sCourseWorkDir, sCourseName)
create_course(psCourseName = sCourseName, psCourseWorkDir = sCourseWorkDir)
```


### A first refinement
The initial version of the function `create_course` as shown above creates a R package for the course. The DESCRIPTION file contains default values that are determined by the devtools specific options stored in the list obtained by `options()`. 

In a first refinement we want to give the user the possibility to specify the information that goes into the DESCRIPTION file more precisely. So let us have a look at the content of the DESCRIPTION file. There is a special function to read the content of DESCRIPTION files called `read.dcf()` where `dcf` stands for Debian Control File. 

```{r}
mDescContent <- read.dcf(file = file.path(sCoursePath, "DESCRIPTION"))
print(mDescContent)
```

The easiest way to specify more information that should go into the description file is to set the devtools specific options, such as shown below.

```{r, eval=FALSE}
## save away old values of options
opar <- options()
## specify new options for devtools
options(devtools.desc.author =  person(given = "Peter", family = "von Rohr", 
                                      email = "peter.vonrohr@gmail.com",
                                      role = c("aut", "cre")), 
        devtools.desc.license = "LGPL-3",
        devtools.name         = "Peter von Rohr",
        devtools.desc         = list(Title       = "Introduction To R", 
                                     Description = "Introductory course to the programming language R"))
## run course creation again
create_course(psCourseName = sCourseName, psCourseWorkDir = sCourseWorkDir)
## restore old options
options(opar)
```

Again looking at the description file of the created course package, we can see that the specified information is included now. 

```{r}
mDescContent <- read.dcf(file = file.path(sCoursePath, "DESCRIPTION"))
print(mDescContent)
```

In principle the above shown solution is fine. The problem is that the dependency between the `RCourseTools` package and 
the `devtools` package is opened to the user. That means, users of the `RCourseTools` package must know about internals 
of the `devtools` package. This is an additional burden that we do not want to impose on the users of the `RCourseTools` 
package. 


### Improved functionality for specifying a DESCRIPTION file
One possibility of specifying information to be written to a DESCRIPTION file is to define a Reference Class (RC) which 
contains the information that we want to be written to the DESCRIPTION file. For more information on reference classes 
and for examples on how to use them, please read the blog post on [Experiments with S4 and Reference Classes](http://charlotte-ngs.github.io/RExperimentsWithS4RC/ExperimentsWithS4AndRef.html). 

### Reference class for information in DESCRIPTION file
When looking at a description file, we recognize that the all fields ot the description file, except the author field 
can be modelled with basic data types. For the author field we are using a separate small reference class which looks 
as follows. 

```{r AuthorRefClass}
#' Reference class to represent an author in a description file
#'
#' @fields   givenName      authors given name
#' @fields   familyName     authors family name
#' @fields   emailAddress   authors email address
#' @fields   role           authors role
DcfAuthorRefClass <- setRefClass(Class = "DcfAuthorRefClass", 
                         fields = list(
                           givenName = "character",
                           familyName = "character",
                           emailAddress = "character",
                           role = "character"
                         ),
                         methods = list(
                           setGivenName = function(psGivenName){
                             "setter for givenName"
                             givenName <<- psGivenName
                           },
                           getGivenName = function(){
                             "getter for givenName"
                             return(givenName)
                           },
                           setFamilyName = function(psFamilyName){
                             "setter for familyName"
                             familyName <<- psFamilyName
                           },
                           getFamilyName = function(){
                             "getter for familyName"
                             return(familyName)
                           },
                           setEmailAddress = function(psEmailAddress){
                             "setter for emailAddress"
                             emailAddress <<- psEmailAddress
                           },
                           getEmailAddress = function(){
                             "getter for emailAddress"
                             return(emailAddress)
                           },
                           setRole = function(psRole){
                             "setter for role"
                             role <<- psRole
                           },
                           getRole = function(){
                             "getter for role"
                             return(role)
                           },
                           addRole = function(psRole){
                             "adding a role to current list of roles"
                             role <<- c(role, psRole)
                           },
                           toDcfString = function(){
                             "writing author information to dcf-formatted string"
                             return(paste("\"", givenName, " ", familyName, 
                                          " <", emailAddress, "> [",
                                          stringi::stri_flatten(role, collapse = ", "),
                                          "]\"", sep = "", collapse = ""))
                           }
                         ))
```

The main reference class for the description file uses the above defined reference class for the author as one component. 

```{r RCDescription}
#' Reference class representing R package description objects
#'
#' @fields title                  dcf field title
#' @fields version                dcf field version
#' @fields author                 dcf field author
#' @fields description            dcf field description
#' @fields depends                dcf field depends
#' @fields licence                dcf field licence
#' @fields LazyData               dcf field LazyData
#' @fields sDcfFileName           output file name
#' @fields mExistingDescription   container for an existing dcf object
RCDesc <- setRefClass(Class = "RCDesc", 
                      fields = list(
                        title = "character",
                        version = "character",
                        author = "DcfAuthorRefClass",
                        description = "character",
                        depends = "character",
                        licence = "character",
                        LazyData = "character",
                        sDcfFileName = "character",
                        mExistingDescription = "matrix"
                      ),
                      methods = list(
                         # setters
                         setTitle = function(pstitle){ title <<- pstitle },
                         setVersion = function(psversion){ version <<- psversion },
                         setAuthor = function(poauthor){ author <<- poauthor },
                         setDescription = function(psdescription){ description <<- psdescription },
                         setDepends = function(psdepends){ depends <<- psdepends },
                         setLicence = function(pslicence){ licence <<- pslicence },
                         setLazyData = function(psLazyData){ LazyData <<- psLazyData },
                         setSDcfFileName = function(pssDcfFileName){ sDcfFileName <<- pssDcfFileName },
                         # getters
                         getTitle = function() { return(title) },
                         getVersion = function() { return(version) },
                         getAuthor = function() { return(author) },
                         getDescription = function() { return(description) },
                         getDepends = function() { return(depends) },
                         getLicence = function() { return(licence) },
                         getLazyData = function() { return(LazyData) },
                         getSDcfFileName = function() { return(sDcfFileName) },
                         # convenience methods for author
                         setGivenName = function(psgivenName){ author$setGivenName(psgivenName) },
                         setFamilyName = function(psfamilyName){ author$setFamilyName(psfamilyName) },
                         setEmailAddress = function(psemailAddress){ author$setEmailAddress(psemailAddress) },
                         setRole = function(psrole){ author$setRole(psrole) },
                         getGivenName = function() { return(author$getGivenName()) },
                         getFamilyName = function() { return(author$getFamilyName()) },
                         getEmailAddress = function() { return(author$getEmailAddress()) },
                         getRole = function() { return(author$getRole()) },
                         # show info
                         show = function() {
                           "Showing current conent of DCF object"
                           cat(" * Title:       ", title, "\n",
                               " * Version:     ", version, "\n",
                               " * Author:      ", author$toDcfString(), "\n",
                               " * Description: ", description, "\n",
                               " * Depends:     ", depends, "\n",
                               " * Licence:     ", licence, "\n",
                               " * LazyData:    ", LazyData, "\n",
                               " * DcfFile:     ", sDcfFileName, "\n"
                               )
                         },
                         # reading an existing description file
                         readDcf = function() {
                           "Reading a description file"
                           # check whether file exists
                           if(file.exists(sDcfFileName)) {
                             mExistingDescription <<- read.dcf(file = sDcfFileName)  
                           } else {
                             cat(" * WARNING: Dcf file: ", sDcfFileName, " NOT FOUND\n")
                           }
                         },
                         # writing the info
                         writeDcf = function() {
                           "Write DCF information to a file"
                             write.dcf(mExistingDescription, file = sDcfFileName)
                         },
                         # add info to existing dcf
                         addToDcf = function() {
                           "Add DCF information to existing description"
                           if(!is.null(title))         mExistingDescription[1,"Title"] <<- title
                           if(length(version) > 0)     mExistingDescription[1,"Version"] <<- version
                           if(length(author) > 0)      mExistingDescription[1,"Author"] <<- author$toDcfString()
                           if(length(description) > 0) mExistingDescription[1,"Description"] <<- description
                           if(length(depends) > 0)     mExistingDescription[1,"Depends"] <<- depends
                           if(length(licence) > 0)     mExistingDescription[1,"License"] <<- licence
                           if(length(LazyData) > 0)    mExistingDescription[1,"LazyData"] <<- LazyData
                         }
                      ))

```

We want to specify the same information with the reference class, as with the `devtools` options above. 

```{r RCDescOption, eval=FALSE}
rCourseDesc <- RCDesc$new()
rCourseDesc$setAuthor(DcfAuthorRefClass$new())
rCourseDesc$setGivenName("Peter")
rCourseDesc$setFamilyName("von Rohr")
rCourseDesc$setEmailAddress("peter.vonrohr@gmail.com")
rCourseDesc$setRole(c("aut","cre"))
rCourseDesc$setLicence("LGPL-3")
rCourseDesc$setTitle("An Introduction To R")
rCourseDesc$setDescription("Introductory course to the programming language R")
rCourseDesc$setSDcfFileName(pssDcfFileName = file.path(sCoursePath, "DESCRIPTION"))
rCourseDesc$show()
rCourseDesc$readDcf()
rCourseDesc$addToDcf()
rCourseDesc$writeDcf()
mDescContent <- read.dcf(file = file.path(sCoursePath, "DESCRIPTION"))
print(mDescContent)
```

The improved functionality of specifying information in the description file is now combined with the creation function.



```{r CreateCourseFinal, eval=FALSE}
#' Create a course with a given name under a given working directory
#'
#' \code{create_course} uses devtools::create to create a package skeleton
#'
#' @param psCourseName      name of the course
#' @param psCourseWorkDir   course working directory
#' @example create_course("MyCourse")
create_course <- function(psCourseName, psCourseWorkDir = getwd(), poDesc = NULL) {
  ##
  ##   create_course(psCourseName, psCourseWorkdir): creates the 
  ##      skeleton of a course in the form of an R package
  ## ########################################################### ##
  ## # check that at least course name argument is specified
  if (is.null(psCourseName)) 
    stop(" * ERROR in create_course: Argument psCourseName missing\n")
  ## # if the parent directory of the course directory does not exist, create it
  if (! dir.exists(psCourseWorkDir)) 
    dir.create(psCourseWorkDir)
  ## # create an R package using devtools::create
  sCoursePath <- file.path(psCourseWorkDir,psCourseName)
  ## # in case old versions exist, remove them
  if (dir.exists(sCoursePath)) unlink(sCoursePath, recursive = TRUE, force = TRUE)
  ## # use devtools to create the package
  devtools::create(path = sCoursePath)
  cat("\n * Created course: ", psCourseName, " in directory: ", psCourseWorkDir, "\n")
  if (!is.null(poDesc)) {
    poDesc$setSDcfFileName(pssDcfFileName = file.path(sCoursePath, "DESCRIPTION"))
    poDesc$readDcf()
    poDesc$addToDcf()
    poDesc$writeDcf()
  }
  ## # copy templates directory
  sTemplateSrcDir <- file.path(system.file(package = "rcoursetools"), "extdata", "templates")
  sTemplateTrgDir <- file.path(sCoursePath, "inst", "extdata")
  if (!dir.exists(sTemplateTrgDir)) dir.create(sTemplateTrgDir, recursive = TRUE)
  cat(" * Copying templates from: ", sTemplateSrcDir, 
      "\n                     to: ", sTemplateTrgDir, "\n")
  file.copy(from = sTemplateSrcDir, to = sTemplateTrgDir, recursive = TRUE, copy.mode = TRUE)
  invisible()
}
```

A course package can now be created with the new function which accepts as additional argument an instance of type `RCDesc`. 

```{r CourseCreateFinal, eval=FALSE}
create_course(psCourseName = sCourseName, psCourseWorkDir = sCourseWorkDir, poDesc = rCourseDesc)
mDescContent <- read.dcf(file = file.path(sCoursePath, "DESCRIPTION"))
print(mDescContent)
```

From the content of the description object printed above, we can see that the information specified in the `RCDesc` reference object is included in the descirption file.


## More Functionality
Once the course skeleton is created, it has to be filled with content. In a first step, we want to provide some tools to create a set of slides to be used in the course.

### Creating Slides
Tools like [`knitr`](http://yihui.name/knitr/) and [`slidify`](http://slidify.org/) help us creating slides that can be used for our course. Package `knitr` is used to compile rnoweb files into pdf documents where `slidify` converts rmarkdown files to html5 pages. 

The function `create_slides()` is a wrapper to both tools `knitr` and `slidify`. 

```{r CreateSlides, eval=FALSE}
#' Create Slides For A Given Course
#'
#' \code{create_slides} creates slides for a given course using a specific slide engine
#'
#' This is a wrapper function that checks whether the required directories exist and 
#' that calls the slide-engine specific functions
#'
#' @param psSlidesName      name of the slides document
#' @param psCourseName      name of the course
#' @param psCourseWorkDir   working directory for course (default getwd())
#' @param psSlideEngine     slide engine, either slidify or knitr
#' @examples
#' create_slides(psCourseName = "rintrocourse", psCourseWorkDir = getwd(), psSlideEngine = "slidify")
create_slides <- function(psSlidesName, psCourseName, psCourseWorkDir, psSlideEngine,
                          psSlidesClass = NULL,
                          plReplace     = NULL){
  ### # check that course work directory exists
  if (! dir.exists(psCourseWorkDir)) dir.create(psCourseWorkDir)
  ### # call slide-engine specific functions
  if (psSlideEngine == "slidify") {
    create_slides_slidify(psSlidesName = psSlidesName, 
                          psCourseName = psCourseName, 
                          psCourseWorkDir = psCourseWorkDir)
  } else if (psSlideEngine == "knitr") {
    create_slides_knitr(psSlidesName    = psSlidesName, 
                        psCourseName    = psCourseName, 
                        psCourseWorkDir = psCourseWorkDir, 
                        psSlidesClass   = psSlidesClass,
                        plReplace       = plReplace)
  } else {
    cat(" *** ERROR in create_slides: slide engine: ", psSlideEngine, " not implemented\n")
  }
}
```

The functions for the specific engines have to be implemented first, before we can create our first set of slides.

```{r CreateSlidesSlidify, eval=FALSE}
#' Create Slides Using Slidify
#'
#' \code{create_slides_slidify} creates a slide document for a given course using slidify
#'
#' @param psSlidesName      name of the slides document
#' @param psCourseName      name of the course
#' @param psCourseWorkDir   working directory for course (default getwd())
create_slides_slidify <- function(psSlidesName, psCourseName, psCourseWorkDir = getwd()){
  ### # put together path for slides
  sSlidesPath <- file.path(psCourseWorkDir, psCourseName, "vignettes", psSlidesName)
  ### # use slidify::author to come up with a skeleton document
  slidify::author(deckdir = sSlidesPath, open_rmd = FALSE)
}
```

The same function for slide-engine `knitr` looks as follows

```{r CreateSlidesKnitr, eval=FALSE}
#' Create a set of slides using knitr as conversion engine
#'
#' The target path where the created set of slides is created
#' is determined by the course work directory, the name of the course
#' and the file name which should be given to the slides. 
#' Based on the slide class name, a set of template files is 
#' chosen wich is copied into the slides path. In the last step
#' the place-holders in the template are replaced by specific 
#' information that is specified. 
create_slides_knitr <- function(psSlidesName, 
                                psCourseName, 
                                psCourseWorkDir, 
                                psSlidesClass = "beamer", 
                                psTemplateDir = file.path(psCourseWorkDir, psCourseName, "inst/extdata/templates"),
                                plReplace     = NULL){
  ### # put together path for slides
  sSlidesPath <- file.path(psCourseWorkDir, psCourseName, "vignettes")
  if (!dir.exists(sSlidesPath)) dir.create(sSlidesPath, recursive = TRUE)
  ### # copy tex files from template directory based on slide class
  copy_template_files(psSlidesPath  = sSlidesPath,
                      psSlidesClass = psSlidesClass,
                      psTemplateDir = psTemplateDir)
  ### # in case a replacement list != null is specified, we do
  ### #  replacement and renaming in on function, otherwise, 
  ### #  we just rename the template
  if (!is.null(plReplace)) {
    ### # replace placeholders with sepcific information
      replace_place_holders(psSlidesName  = psSlidesName, 
                            psSlidesPath  = sSlidesPath,
                            psSlidesClass = psSlidesClass,
                            plReplace     = plReplace)    
  } else {
    ### # no replacement list specified, just
    ### #  rename template source files in copied slide-path
    rename_template_files(psSlidesName  = psSlidesName,
                          psSlidesPath  = sSlidesPath,
                          psSlidesClass = psSlidesClass)
    
  }
  cat(" * Set of knitr slides created under: ", sSlidesPath, "\n",
      " * Slides can be modified by opening project: ", psCourseName, " and running \n",
      "   > file.edit(\"vignettes/", paste(psSlidesName, ".rnw", sep = ""), 
      "\")\n", sep = "")

  invisible()
}

#' Copy template source and class files based on psSlidesClass to slides directory
#'
#' @param psSlidesPath    target path of slides source file 
#' @param psSlidesClass   latex class of slides
#' @param psTemplateDir   template directory where templates of sources are stored
copy_template_files <- function(psSlidesPath, psSlidesClass, psTemplateDir) {
  sTemplatePath <- file.path(psTemplateDir, psSlidesClass)
  sTemplateFiles <- list.files(path = sTemplatePath)
  for (f in sTemplateFiles) {
    file.copy(from = file.path(sTemplatePath, f ), to = psSlidesPath, recursive = TRUE, copy.mode = TRUE)
  }
    
}


#' Renames the tex/rnw source template files to match the specified file name for the slides
#'
#' @param psSlidesName    the final file name of the slides
#' @param psSlidesPath    path to the slides target directory
#' @param psSlidesClass   name of the utilsed latex document class
rename_template_files <- function(psSlidesName, psSlidesPath, psSlidesClass) {
  file.rename(from = file.path(psSlidesPath, paste(psSlidesClass, ".rnw", sep = "")), 
              to   = file.path(psSlidesPath, paste(psSlidesName, ".rnw", sep = "")))
}

#' Replace placeholders with sepcific information given in replacement list
#'
#' Template source files contain placeholders of the format [REPLACE_WITH_<TAG>] where 
#' tag is a unique identifier that is used as name in the replacement list. The 
#' replacement list has the tags as names and as values the string that should  
#' replace the placeholder. In a loop over the replacement list entries, the 
#' placeholders are constructed from the tags which are the names in the replacement 
#' list and the placeholders are replaced by the values in the replacement list.
#'
#' @param psSlidesName    Name of the slides document
#' @param psSlidesPath    Path to where the slides are stored
#' @param psSlidesClass   Name of the latex document class used for the slides
#' @param plReplace       replacement list that maps placeholders to values in the final slides document
replace_place_holders <- function(psSlidesName, psSlidesPath, psSlidesClass, plReplace) {
  ### # read slides source file into character vector
  sSlideSourceFile <- file.path(psSlidesPath, paste(psSlidesClass, ".rnw", sep = ""))
  conSlideSource <- file(description = sSlideSourceFile)
  sSlidesSource <- readLines(con = conSlideSource)
  close(con = conSlideSource)
  ### # original template file is not needed anymore
  file.remove(sSlideSourceFile)
  
  ### # loop over entries in replacement list and do the substitution
  for (sCurReplName in names(plReplace)) {
    sCurPat <- paste("[REPLACE_WITH_", sCurReplName, "]", sep = "", collapse = "")
    sSlidesSource <- gsub(sCurPat, plReplace[[sCurReplName]], sSlidesSource, fixed = TRUE)
  }
  
  ### # output file
  sOutFile <- file.path(psSlidesPath, paste(psSlidesName, ".rnw", sep = ""))
  cat(sSlidesSource, file = sOutFile, sep = "\n")
  cat(" * Placeholders replaced\n")
  invisible()
}

```

Now we are ready to create our first set of slides. 

```{r IntroSlidesSlidify, eval = FALSE}
create_slides(psSlidesName = "IntroductionToR", 
              psCourseName = sCourseName, 
              psCourseWorkDir = sCourseWorkDir, 
              psSlideEngine = "slidify")
```

